### 2022-11-28 Paper2.Rmd
commit

---
title: "Paper 2, Econ 980x"
author: "Alice Chen"
date: "2022-11-27"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(ggplot2)
library(janitor)
library(ggthemes)
library(stargazer)

```

```{r files}

csv <- read_csv(file = "atus_00001.csv.gz") %>%
  clean_names()

# race 
race_file <- read_excel("race_980x.xlsx")

race_file <- race_file %>%
  rename(code1 = race) %>%
  rename(race = code)

csv <- left_join(csv, race_file, by = "race") %>%
  rename(race_code = race) %>%
  rename(race = code1)

# education
education_file <- read_excel("Education.xlsx")
education_file <- education_file %>%
  rename(code_educ = educ) %>%
  rename(educ = code_education) %>%
  mutate(educ = as.numeric(educ))

csv <- left_join(csv, education_file, by = "educ") %>%
  rename(code_education = educ) %>%
  rename(educ = code_educ)

# marital status
marital_file <- read_excel("marital_status.xlsx")
education_file <- marital_file %>%
  mutate(marst = as.numeric(marst))

csv <- left_join(csv, marital_file, by = "marst") %>%
  rename(code_marst = marst) %>%
  rename(marst = marst_code) %>%
  mutate(marst_simple = ifelse(code_marst <= 2, "Married", ifelse(code_marst >= 6, "Never Married", "Separated/Divorced")))

# age
csv <- csv %>%
  filter(age > 18)

# children
csv <- csv %>%
  mutate(child = ifelse(yngch < 19, 1, 0) )

# gender

csv <- csv %>%
  mutate(gender = ifelse(sex == 1, "male", "female"))

# years

```

```{r graphs}

#leisure hours aggregate
csv %>%
  ggplot(mapping = aes(x = bls_leis)) +
  geom_histogram(binwidth=50) +
  labs(
  title = "Distribution of annual individual leisure hours",
  x = "Annual Hours",
  y = "Count",
  caption = "Data from ATUS",
)

csv %>%
  filter(year %in% c(2007, 2009)) %>%
  ggplot(mapping = aes(x = bls_leis)) +
  geom_histogram(binwidth=50) +
  facet_wrap(~year) +
  labs(
  title = "Distribution of annual individual leisure hours",
  x = "Annual Hours",
  y = "Count",
  caption = "Data from ATUS",
)

#leisure hours aggregate by gender
csv %>%
  group_by(gender) %>%
  ggplot(mapping = aes(x = bls_leis) )+
  geom_histogram(binwidth=50) +
  facet_wrap(~ gender) +
    labs(
  title = "Distribution of annual individual leisure hours by gender",
  x = "Annual Hours",
  y = "Count",
  caption = "Data from ATUS",
)

csv %>%
  filter(year %in% c(2007, 2009)) %>%
  ggplot(mapping = aes(x = bls_leis)) +
  geom_histogram(binwidth=50) +
  facet_wrap(~ gender + year) +
  labs(
  title = "Distribution of annual individual leisure hours by gender",
  x = "Annual Hours",
  y = "Count",
  caption = "Data from ATUS",
)

# Median
csv %>%
  group_by(year, gender) %>%
  summarize(median(bls_leis)) %>%
  clean_names() %>%
  ggplot(aes(x = year, y = median_bls_leis)) +
  geom_line() +
  facet_wrap(~ gender) +
  labs(
  title = "Median annual individual leisure hours by gender",
  y = "Annual Hours",
  x = "Year",
  caption = "Data from ATUS",
)

# All Quantile
csv %>%
  group_by(year, gender) %>%
  summarize(quantile_25 = quantile(bls_leis, 0.25), quantile_50 = quantile(bls_leis, 0.5), quantile_75 = quantile(bls_leis, 0.75)) %>%
  clean_names() %>%
  pivot_longer(quantile_25:quantile_75, names_to = "quantile", values_to = "hours") %>%
  ggplot(aes(x = year, y = hours, group = quantile, colour = quantile)) +
  geom_line() +
    facet_wrap(~ gender) +
  labs(
  title = "Annual individual leisure hours by gender",
  y = "Annual Hours",
  x = "Year",
  caption = "Data from ATUS",
)

csv %>%
  group_by(year) %>%
  summarize(quantile(bls_leis, 0.25), quantile(bls_leis, 0.5), quantile(bls_leis, 0.75)) %>%
  clean_names() %>%
  pivot_longer(!year, names_to = "quantile", values_to = "hours") %>%
  ggplot(aes(x = year, y = hours, group = quantile)) +
  geom_line()

csv %>%
  group_by(year) %>%
  summarize(quantile(bls_leis, 0.25), quantile(bls_leis, 0.5), quantile(bls_leis, 0.75)) %>%
  clean_names() %>%
  pivot_longer(!year, names_to = "quantile", values_to = "hours") %>%
  filter(year > 2006) %>%
  filter(year < 2011) %>%
  ggplot(aes(x = year, y = hours, group = quantile)) +
  geom_line()

# by year, gender
csv %>%
  group_by(year, gender) %>%
  summarize(quantile_25 = quantile(bls_leis, 0.25), quantile_50 = quantile(bls_leis, 0.5), quantile_75 = quantile(bls_leis, 0.75)) %>%
  clean_names() %>%
  pivot_longer(quantile_25:quantile_75, names_to = "quantile", values_to = "hours") %>%
  filter(year > 2005) %>%
  filter(year < 2011) %>%
  ggplot(aes(x = year, y = hours, group = quantile, colour = quantile)) +
  geom_line() +
  facet_wrap(~gender) +
    labs(
  title = "Annual individual leisure hours by gender",
  y = "Annual Hours",
  x = "Year",
  caption = "Data from ATUS",
)

# by year, gender
csv %>%
  group_by(year, gender) %>%
  summarize(median = median(bls_leis)) %>%
  clean_names() %>%
  filter(year > 2005) %>%
  filter(year < 2011) %>%
  ggplot(aes(x = year, y = median)) +
  geom_line() +
  facet_wrap(~gender) +
      labs(
  title = "Annual individual leisure hours by gender",
  y = "Annual Hours",
  x = "Year",
  caption = "Data from ATUS",
)

#csv %>%
 # ggplot(mapping = aes(x = year, y = bls_leis)) + 
 # geom_jitter(alpha = 0.01)

```

```{r regressions, echo=FALSE}

#naive regression of men versus women

g <- csv %>%
  lm(formula = bls_leis ~ gender)

#naive regression of men versus women + year

y <- csv %>%
  lm(formula = bls_leis ~ gender + year)

# race count
newcsv <- csv %>%
 filter(race_code <= 210)

newcsv
newcsv %>%
  lm(formula = bls_leis ~ gender + year + age + race)

# 
a <- newcsv %>%
  lm(formula = bls_leis ~ gender + year + age + race) 
summary(a)

newcsv

# education
newcsv1 <- newcsv %>%
  mutate(new_educ = ifelse(code_education <= 17, "<High School", ifelse(code_education >= 40, ">Four Year Degree", "High School Graduate")))
  newcsv1

b <- newcsv1 %>%
  lm(formula = bls_leis ~ gender + year + age + race + new_educ) 
summary(b)

newcsv1

# with children

c <- newcsv1 %>%
  lm(formula = bls_leis ~ gender + year + age + race + new_educ + nchild) 
summary(c)

#marital status

d <- newcsv1 %>%
  lm(formula = bls_leis ~ gender + year + age + race + new_educ + nchild + marst_simple) 
summary(d)

stargazer(g, y, a, title="Various Models", align=TRUE)
```

```{r moremodels}
#Change

change <- newcsv1 %>%
  group_by(gender,  year, new_educ, child, marst_simple) %>%
  summarize(median(bls_leis)) %>%
  clean_names() %>%
  pivot_wider(values_from = median_bls_leis, names_from = year) %>%
  clean_names() %>%
  mutate(leisure_change = x2009 - x2007)

change2 <- newcsv1 %>%
  group_by(gender, year) %>%
  summarize(median(bls_leis)) %>%
  clean_names() %>%
  pivot_wider(values_from = median_bls_leis, names_from = year) %>%
  clean_names() %>%
  mutate(leisure_change = x2009 - x2007)

change2

delta <- change %>%
  lm(formula = leisure_change ~ gender+ new_educ+ child+ marst_simple)
summary(delta)

```